---
title: "Project Introduction"
author: "Ido Shachar & Tal Kozlosvki"
output:
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r}
source('R\\help_functions.r')
```



```{r}
data <- read.csv('data\\data_20190828.csv')


data %<>% 
  dplyr::select(-term, -weight, 
                -APGAR1, -APGAR5, 
                -mat_fever, -risk_at_birth, 
                -risk_clinical, 
                -broad_ABX_moreThan4h, 
                -broad_ABX2_3.9h, 
                -GBS_ABX_moreThan2h, 
                -ABX_lessThan2h, 
                -risk_clinical, 
                -rec_none, -rec_culture, 
                -rec_ABX, 
                -fever, 
                -hypothermia, 
                -resp_distress, 
                -maternal_reason, 
                -Other, -total_taken, 
                -X2nd_value, 
                -hour2, 
                -X3rd_value, 
                -hour3, -peak_CRP, 
                -high_CRP, 
                -symp_lessThan12h, 
                -symp12_24h, 
                -symp_moreThan24h, 
                -LP, 
                -ABX, 
                -ABX_dur, 
                -NICU, 
                -pos_blood_culture, 
                -other_culture, 
                -contamin, 
                -high_suspicion, 
                -no_XR, 
                -jaundice_workup, 
                -Nati, -Sivan, 
                -GBS_neg, 
                -GBS_unk, 
                -pneumonia, 
                -PTX, 
                -phototherapy
          ) %>%
  filter(ROMtime < 90, !is.na(hour1)) %>% 
  mutate("NVD" = ifelse(VE == 1, 1, NVD), 
         "log_CRP" = log(X1st_value + 0.1),
         # "CRP_trans" = X1st_value ^ (1/3),
         "log_hour" = log(hour1), 
         "T_grp" = ifelse(NVD == 1, "regular", 
                          ifelse(CS_elect == 1, "CS_elect", "CS_emerg"))) %>% 
  dplyr::select(-VE, -ID, -AGA, -SGA, -NVD, -CS_elect, -CS_emerg, 
                -PROM, -IAP_A, -IAP_I, -no_symp, -X1st_value, -hour1, -log_hour)


str(data)

```


```{r}
dim(data)
table(data$T_grp)
```


```{r}

data %>% 
  group_by(T_grp) %>% 
  summarise_all(.funs = c("mean" = mean, 
                          "sd" = sd)) %>% 
  pivot_longer(cols = c(-T_grp)) %>% 
  mutate("stat" = ifelse(grepl("_mean", name), "mean", "sd")) %>%
  mutate("name" = gsub("_mean|_sd", "", name)) %>%
  pivot_wider(id_cols = name, 
              names_from = c(T_grp, stat), 
              values_from = value) %>% 
  write.csv('output\\desc_stats.csv', row.names = F)

```


```{r}
data %>% 
  group_by(T_grp) %>% 
  summarise("perc_Female" = sum(gender) / n(), 
            "perc_preterm" = sum(preterm) / n(),
            "mean_CRP" = mean(log_CRP), 
            "sd_CRP" = sd(log_CRP), 
            "mean_gestAge" = mean(gest_age), 
            "sd_gestAge" = sd(gest_age),
            "perc_MSAF" = sum(MSAF) / n(), 
            "perc_DM" = sum(DM) / n(), 
            "mean_ROMtime" = mean(ROMtime), 
            "sd_ROMtime" = sd(ROMtime),
            "perc_lGA" = sum(LGA) / n(),
            "n" = n()) %>% 
  knitr::kable()
  # write.csv('dem.csv') 



```


### Plot descriptive statistics


```{r}
data %>% 
  dplyr::select(T_grp, gest_age, ROMtime, high_mat_temp) %>% 
  pivot_longer(cols = c(gest_age, ROMtime, high_mat_temp)) %>% 
  ggplot(aes(T_grp, value, fill = T_grp)) + 
  geom_violin(alpha = 0.5) + 
  geom_point() + 
  facet_grid(name ~. , scales = "free")

#### Selected plot
data %>% 
  dplyr::select(T_grp, gest_age, ROMtime, high_mat_temp) %>% 
  mutate("log_ROM" = log(ROMtime + 0.1)
         # , 
         # "log_mat_temp" = log(high_mat_temp + 0.1)
         ) %>%
  pivot_longer(cols = c(gest_age, log_ROM, high_mat_temp)) %>% 
    ggplot(aes(T_grp, value, fill = T_grp)) + 
  geom_violin(alpha = 1) + 
  geom_point() + 
  facet_grid(name ~. , scales = "free")
  


data %>% 
  dplyr::select(T_grp, gest_age, ROMtime, high_mat_temp) %>% 
  mutate("log_ROM" = log(ROMtime + 0.1), 
         "log_mat_temp" = log(high_mat_temp + 0.1)) %>%
  pivot_longer(cols = c(gest_age, log_ROM, log_mat_temp)) %>% 
  ggplot(aes(value, fill = T_grp)) + 
  geom_histogram(alpha = 0.3) + 
  facet_grid(~ name, scales = "free")

data %>%
  ggplot(aes(T_grp, log_CRP)) + 
  geom_boxplot() + 
  geom_point()


data %>% 
  mutate("Group" = ifelse(T_grp == "CS_elect", "CS elective", 
                          ifelse(T_grp == "CS_emerg", 
                                 "CS urgent", "Regular birth"))) %>% 
  ggplot(aes(Group, log_CRP, fill = Group)) + 
  geom_violin(alpha = 0.3) + 
  geom_point() + 
  theme_bw() +
  ggtitle('log CRP per Treatment Group') + 
  scale_fill_discrete(name = "Group") + 
  xlab("") + 
  ylab("log CRP") +
  theme(legend.position = "none")


data %>% 
  ggplot(aes(fill = T_grp, log_CRP)) + 
  geom_density(alpha = 0.15)

```




```{r}
data_elective <- data %>% 
  filter(T_grp != 'CS_emerg') %>% 
  mutate("T_grp" = ifelse(T_grp == "regular", 0, 1)) %>% 
  dplyr::select(-postterm)

prop_elective <- prop_score_fun(data = data_elective, TreatCol = "T_grp", OutcomeCol = "log_CRP")

  
data_elective %>% 
  mutate("prop" = prop_elective) %>% 
  ggplot(aes(prop, fill = as.factor(T_grp))) + geom_histogram(alpha = 0.3)

# data_elective %>% 
#   mutate("prop" = prop_elective) %>% 
#   ggplot(aes(prop, x = as.factor(T_grp))) + 
#   geom_boxplot()

data_elective %>% 
  mutate("prop" = prop_elective) %>% 
  group_by(T_grp) %>% 
  summarise(sum(prop > 0.7), n())


match_att_elective <- matching_att_fun(data = data_elective, 
                               TreatCol = "T_grp", 
                               OutcomeCol = "log_CRP", 
                              thresh = 1)

plot_data <- match_att_elective$data_plot
plot_data %>% 
  mutate("Group" = ifelse(T_grp == 0, "Regular", "CS_elect")) %>%
  ggplot(aes(x = Group, y = log_CRP)) + 
  geom_boxplot() + 
  ggtitle("Elective")



T.att_elective <- Tlearner_att_fun(data = data_elective, 
                           TreatCol = "T_grp", 
                           OutcomeCol = "log_CRP" )

S.att_elective <- Slearner_att_fun(data = data_elective, 
                           TreatCol = "T_grp", 
                           OutcomeCol = "log_CRP")
match_att_elective
print("T.att_elective")
T.att_elective
print("S.att_elective")
S.att_elective
```



```{r}
### Sensitivity analysis
sendata <- matrix(plot_data$log_CRP, ncol = 2)
senmw(sendata, gamma = 1, method = "t")$pval
senmw(sendata, gamma = 1.1, method = "t")$pval
senmw(sendata, gamma = 1.2, method = "t")$pval
senmw(sendata, gamma = 1.3, method = "t")$pval

senmwCI(sendata, gamma = 1, method = "t", one.sided = TRUE)
senmwCI(sendata, gamma = 1.1, method = "t", one.sided = TRUE)
senmwCI(sendata, gamma = 1.2, method = "t", one.sided = TRUE)
senmwCI(sendata, gamma = 1.3, method = "t", one.sided = TRUE)
```


```{r}
#### Descriptive statistics table 

elective_desc <- data_elective %>% 
  group_by(T_grp) %>% 
  summarise_all(.funs = c("Mean" = mean, "SD" = sd)) %>% 
  pivot_longer( cols = -T_grp, names_to = "Variable_full") %>% 
  mutate("fun" = str_replace(Variable_full, "^.+_", "")) %>% 
  rowwise() %>%
  mutate("Variable" = gsub(paste0("_", fun), "", Variable_full)) %>% 
  dplyr::select(-Variable_full) %>% 
  pivot_wider(id_cols = Variable, names_from = c(T_grp, fun), values_from = value) %>% 
  mutate("Pvalue_elect" = NA) %>%
  as.data.frame()


reg.ins <- which(data_elective$T_grp == 0)
ele.ins <- which(data_elective$T_grp == 1)

for(i in 1:nrow(elective_desc)) {
  tmp_var <- elective_desc[i, "Variable"]
  if(length(unique(data_elective[, tmp_var])) == 2) {
        tmp.tab <- table(data_elective[, "T_grp"], 
                     data_elective[, tmp_var])
    if(any(tmp.tab <= 5)) {
      elective_desc[i, "Pvalue_elect"] <- fisher.test(tmp.tab)$p.value
    } else {
     elective_desc[i, "Pvalue_elect"] <- chisq.test(tmp.tab)$p.value
    }
      
  } else {
    elective_desc[i, "Pvalue_elect"] <- t.test(data_elective[reg.ins, tmp_var], 
                                         data_elective[ele.ins, tmp_var])$p.value
  }
}



```





### Elective - filtered data (prop < 0.7)

```{r}

match_att_electf <- matching_att_fun(data = data_elective, 
                               TreatCol = "T_grp", 
                               OutcomeCol = "log_CRP", 
                              thresh = 0.7)

prop_elective <- prop_score_fun(data = data_elective, TreatCol = "T_grp", OutcomeCol = "log_CRP")

data_elective_filter <- data_elective %>% 
  mutate("prop" = prop_elective) %>% 
  filter(prop < 0.7) %>% 
  dplyr::select(-prop)

plot_data <- match_att_electf$data_plot
plot_data %>% 
  mutate("Group" = ifelse(T_grp == 0, "Regular", "CS_elect")) %>%
  ggplot(aes(x = Group, y = log_CRP)) + 
  geom_boxplot() + 
  ggtitle("Filtered Elective")

T.att_electf <- Tlearner_att_fun(data = data_elective_filter, 
                           TreatCol = "T_grp", 
                           OutcomeCol = "log_CRP" )

S.att_electf <- Slearner_att_fun(data = data_elective_filter, 
                           TreatCol = "T_grp", 
                           OutcomeCol = "log_CRP")
match_att_electf
print("T.att_electf")
T.att_electf
print("S.att_electf")
S.att_electf
```



```{r}
### Sensitivity analysis
sendata <- matrix(plot_data$log_CRP, ncol = 2)
senmw(sendata, gamma = 1, method = "t")$pval
senmw(sendata, gamma = 1.5, method = "t")$pval
senmw(sendata, gamma = 1.6, method = "t")$pval
senmw(sendata, gamma = 1.75, method = "t")$pval

senmwCI(sendata, gamma = 1, method = "t", one.sided = TRUE)
senmwCI(sendata, gamma = 1.5, method = "t", one.sided = TRUE)
senmwCI(sendata, gamma = 1.6, method = "t", one.sided = TRUE)
senmwCI(sendata, gamma = 1.75, method = "t", one.sided = TRUE)
```

#### Emergency 





```{r}
data_urgent <- data %>% 
  filter(T_grp != 'CS_elect') %>% 
  mutate("T_grp" = ifelse(T_grp == "regular", 0, 1)) 

prop_urgent <- prop_score_fun(data = data_urgent, 
                                TreatCol = "T_grp", 
                                OutcomeCol = "log_CRP")

  
data_urgent %>% 
  mutate("prop" = prop_urgent) %>% 
  ggplot(aes(prop, fill = as.factor(T_grp))) + geom_histogram(alpha = 0.3)

data_urgent %>% 
  mutate("prop" = prop_urgent) %>% 
  group_by(T_grp) %>% 
  summarise(sum(prop > 0.55), n())


match_att_urgent <- matching_att_fun(data = data_urgent, 
                               TreatCol = "T_grp", 
                               OutcomeCol = "log_CRP", 
                              thresh = 0.55)

plot_data <- match_att_urgent$data_plot
plot_data %>% 
  mutate("Group" = ifelse(T_grp == 0, "Regular", "CS_emerg")) %>%
  ggplot(aes(x = Group, y = log_CRP)) + 
  geom_boxplot() + 
  ggtitle("Cesarean Emergency")


T.att_urgent <- Tlearner_att_fun(data = data_urgent, 
                           TreatCol = "T_grp", 
                           OutcomeCol = "log_CRP" )

S.att_urgent <- Slearner_att_fun(data = data_urgent, 
                           TreatCol = "T_grp", 
                           OutcomeCol = "log_CRP")
match_att_urgent
print("T.att_urgent")
T.att_urgent
print("S.att_urgent")
S.att_urgent
```


```{r}
### Sensitivity analysis
sendata <- matrix(plot_data$log_CRP, ncol = 2)
senmw(sendata, gamma = 1, method = "t")$pval
# senmw(sendata, gamma = 1.5, method = "t")$pval
# senmw(sendata, gamma = 1.6, method = "t")$pval
# senmw(sendata, gamma = 1.75, method = "t")$pval

senmwCI(sendata, gamma = 1, method = "t", one.sided = TRUE)
# senmwCI(sendata, gamma = 1.5, method = "t", one.sided = TRUE)
# senmwCI(sendata, gamma = 1.6, method = "t", one.sided = TRUE)
# senmwCI(sendata, gamma = 1.75, method = "t", one.sided = TRUE)
```



```{r}
#### Descriptive statistics table 

 
urgent_desc <- data_urgent %>% 
  group_by(T_grp) %>% 
  summarise_all(.funs = c("Mean" = mean, "SD" = sd)) %>% 
  pivot_longer( cols = -T_grp, names_to = "Variable_full") %>% 
  mutate("fun" = str_replace(Variable_full, "^.+_", "")) %>% 
  rowwise() %>%
  mutate("Variable" = gsub(paste0("_", fun), "", Variable_full)) %>% 
  dplyr::select(-Variable_full) %>% 
  pivot_wider(id_cols = Variable, names_from = c(T_grp, fun), values_from = value) %>% 
  mutate("Pvalue_urgent" = NA) %>%
  as.data.frame()


reg.ins <- which(data_urgent$T_grp == 0)
urg.ins <- which(data_urgent$T_grp == 1)

for(i in 1:nrow(urgent_desc)) {
  tmp_var <- urgent_desc[i, "Variable"]
  if(length(unique(data_urgent[, tmp_var])) == 2) {
    tmp.tab <- table(data_urgent[, "T_grp"], 
                     data_urgent[, tmp_var])
    if(any(tmp.tab <= 5)) {
      urgent_desc[i, "Pvalue_urgent"] <- fisher.test(tmp.tab)$p.value
    } else {
     urgent_desc[i, "Pvalue_urgent"] <- chisq.test(tmp.tab)$p.value
    }

  } else {
    urgent_desc[i, "Pvalue_urgent"] <- t.test(data_urgent[reg.ins, tmp_var], 
                                         data_urgent[urg.ins, tmp_var])$p.value
  }
}

colnames(elective_desc)[c(4:5)] <- paste0("elective", "_", colnames(elective_desc)[c(4:5)])
colnames(urgent_desc)[c(4:5)] <- paste0("urgent", "_", colnames(urgent_desc)[c(4:5)])

full_desc <- urgent_desc %>% 
  full_join(elective_desc)

write.csv(full_desc, 'Output\\descritpive_table.csv', row.names = F)
```





#### Emergency - filtered (prop < 0.55)

```{r}


prop_urgent <- prop_score_fun(data = data_urgent, 
                              TreatCol = "T_grp", 
                              OutcomeCol = "log_CRP")

data_urgent_filter <- data_urgent %>% 
  mutate("prop" = prop_urgent) %>% 
  filter(prop < 0.55) %>% 
  dplyr::select(-prop)


match_att_urgentf <- matching_att_fun(data = data_urgent_filter, 
                               TreatCol = "T_grp", 
                               OutcomeCol = "log_CRP", 
                              thresh = 1)

plot_data <- match_att_urgentf$data_plot
plot_data %>% 
  mutate("Group" = ifelse(T_grp == 0, "Regular", "CS_emerg")) %>%
  ggplot(aes(x = Group, y = log_CRP)) + 
  geom_boxplot() + 
  ggtitle("Cesarean Emergency Filtered")


T.att_urgentf <- Tlearner_att_fun(data = data_urgent_filter, 
                           TreatCol = "T_grp", 
                           OutcomeCol = "log_CRP" )

S.att_urgentf <- Slearner_att_fun(data = data_urgent_filter, 
                           TreatCol = "T_grp", 
                           OutcomeCol = "log_CRP")
match_att_urgentf
print("T.att_urgentf")
T.att_urgentf
print("S.att_urgentf")
S.att_urgentf
```



```{r}

### Sensitivity analysis
sendata <- matrix(plot_data$log_CRP, ncol = 2)
senmw(sendata, gamma = 1, method = "t")$pval
# senmw(sendata, gamma = 1.5, method = "t")$pval
# senmw(sendata, gamma = 1.6, method = "t")$pval
# senmw(sendata, gamma = 1.75, method = "t")$pval

senmwCI(sendata, gamma = 1, method = "t", one.sided = TRUE)
# senmwCI(sendata, gamma = 1.5, method = "t", one.sided = TRUE)
# senmwCI(sendata, gamma = 1.6, method = "t", one.sided = TRUE)
# senmwCI(sendata, gamma = 1.75, method = "t", one.sided = TRUE)
```
